---
title: "Developing Data Products - Week 3 Project"
author: "Kurt Fitz"
date: "15/02/2021"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r ECHO=FALSE, warning=FALSE, message=FALSE}
library(plotly)
library(WDI)

```

## 1. GDP Data for Canada, Mexico, and U.S.A

```{r ECHO=FALSE, warning=FALSE, message=FALSE}
dat = WDI(indicator='NY.GDP.PCAP.KD', country=c('MX','CA','US'), start=1960, end=2012)

plot_ly(dat, x=dat$country, y=dat$NY.GDP.PCAP.KD, mode="markers")

```

## 2. GDP Data for Canada, Mexico, and U.S.A
```{r ECHO=FALSE, warning=FALSE, message=FALSE}

plot_ly(dat, x=dat$year, y=dat$NY.GDP.PCAP.KD, mode="markers", color=as.factor(dat$country))

```

## Final Slide

Review Criteria
The rubric contains the following two questions:

1. Does the web page feature a date and is this date less than two months before the date that you're grading this assignment?
2. Is the web page a presentation and does it feature an interactive plot that appears to have been created with Plotly?


