---
title: "Statistical Inference Course Project"
author: "Kurt Fitz"
date: "April 21, 2019"
output:
pdf_document: default
html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

#Overview

The following document explore some basic inference using the ToothGrowth data in R, and is the second half the two part course project assignment.


## Part 2 Basic Inferential Data Analysis of ToothGrowth data in R datasets package.
The following basic inferential analysis will compare tooth growth by dose and supplement using confidence intervals.

Below is a basic summary of the tooth growth data:
```{r}
library(datasets)
data("ToothGrowth")

head(ToothGrowth)
dim(ToothGrowth)
str(ToothGrowth)
summary(ToothGrowth)
```

## Tooth Growth by Supplement
```{r}
library(ggplot2)

qplot(supp, len, data=ToothGrowth, facets=~dose, main="Tooth growth by supplement type and dosage (mg)", xlab="Supplement", ylab="Tooth length") + geom_boxplot(aes(fill = supp))

t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = ToothGrowth)

```

We can see from the Student's t-test that the 95% confidence interval contains 0; therefore, we cannot reject the null hypotheses that there is no effect between the two supplements.

## Tooth Growth by Dose
Next, we can examine the relationship between tooth growth and dosage.
```{r}

## Subset to dosage 0.5 - 1.0
doseRange1 <- subset(ToothGrowth, dose %in% c(0.5, 1.0))

## Subset to dosage 0.5 - 2.0
doseRange2 <- subset(ToothGrowth, dose %in% c(0.5, 2.0))

## Subset to dosage 1.0 - 2.0
doseRange3 <- subset(ToothGrowth, dose %in% c(1.0, 2.0))

t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = doseRange1)

```

```{r}
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = doseRange2)
```

```{r}
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = doseRange3)
```

## Conclusions
From the preceeding t-tests performed on the three dose ranges, we can conclude that there is a relationship between the increase and dosage and increase in growth. The p-value for each of the three ranges was below 0 and the 95% confidence intervals do not contain 0.
